---
title: "RMarkdown Tutorial"
author: "Hao Ran Lai"
date: "16 March 2020"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(DiagrammeR)
library(DiagrammeRsvg)
library(magrittr)
library(rsvg)
```

```{r echo=FALSE}
knitr::knit_hooks$set(mysize = function(before, options, envir) {
  if (before) 
    return(options$size)
})
```

## R Markdown

Combining scientific analyses and writing in one programme.

```{r}
grViz("digraph flowchart {
      # node definitions with substituted label text
      node [fontname = Helvetica, shape = rectangle]        
      tab1 [label = '@@1']
      tab2 [label = '@@2']
      tab3 [label = '@@3']
      tab4 [label = '@@4']
      tab5 [label = '@@5']
      tab6 [label = '@@6']

      # edge definitions with the node IDs
      tab1 -> tab3;
      tab2 -> tab3;
      tab3 -> tab4 -> tab5 -> tab6
      }

      [1]: 'Analyses in R'
      [2]: 'Writing in Markdown'
      [3]: 'Report / assignment in RMarkdown'
      [4]: 'Knit via knitr'
      [5]: 'Document conversion via pander'
      [6]: 'Final form in PDF, HTML, Word...'
      ",
      engine = "circo") %>%
    export_svg %>% charToRaw %>% rsvg_pdf("workflow.pdf")
```

```{r out.height="3in", fig.align='center'}
knitr::include_graphics("workflow.pdf")
```

## Installation

```
install.packages("rmarkdown")
```

- Or just install RStudio, which comes with RMarkdown thesedays.
- To compile PDFs, you may need to manually:

```
install.packages(“tinytex”)
tinytex::install_tinytex()
```

## How to learn RMarkdown

- Come to this session (peer support)
- RMarkdown cheatsheets: [link](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf)
- RStudio > Help > Markdown Quick Reference
- Google

Some examples: [link](https://rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)

## Why all the fuss?

- Reproducibility

```{r fig.height=4, fig.width=5, fig.align='center', out.width = '70%'}
x <- seq(0, 1, length.out = 100)
y1 <- exp(-10 * x - 6 * x^2)
y2 <- exp(-4 * x + 4* x^2 - 1.3* x^3)
par(mar = c(5,4,1,1))
plot(0, 0, type = "n", xlim = c(0, 1), ylim = c(0, 1), xaxt = "n", yaxt = "n", bty = "l",
     xlab = "Frequency of boss/collaborator asking to make changes",
     ylab = "Efficiency / Reproducibility")
axis(1, c(0.1, 0.9), c("Just doesn't care", "Like your mom"), tick = FALSE)
lines(x, y1, col = "red", lwd = 2)
lines(x, y2, col = "blue", lwd = 2)
legend(0.5, 1, lty=c(1,1), col=c("blue","red"), legend=c("RMarkdown","R + MS Word"), lwd=c(2,2))
```

- More of a "one-stop-shop" than \LaTeX

## Reproducible plot

```{r echo=TRUE, fig.height=3.5, fig.width=4}
x <- seq(0, 1, length.out = 100)
y <- x + rnorm(length(x), 0, 0.1)
plot(x, y)
```

## Reproducible analyses

```{r echo=TRUE, mysize=TRUE, size = '\\footnotesize'}
mod <- lm(y ~ x)
summary(mod)
```

## Reproducible analyses

```{r echo=TRUE, fig.height=3.5, fig.width=4}
plot(x, y)
abline(mod, col = "red", lwd = 2)
```

